---
title: "DDS Case Study 1"
author: "Armando Vela & Paul Adams"
date: "June 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## P&A Consulting Group
####  Budweiser Inc. Beer & Brewery Analysis

After carefully reviewing the beer offerings data and brewery data provided by your company, our consulting group was able to uncover helpful information in choosing which beer type to offer within specific US states. Our analysis, listed below, mainly dealt with reviewing the alcohol content (ABV) and international bitterness unit (IBU) for each beer type within each state.  

Some findings we identified were that if your company hoped to offer a more bitter beer (higher IBU), than the state of West Virginia should be used for the initial roll-out since the citizens are already accustomed to a more bitter beer since it has the highest median IBU value in the US.  Also, if you were going to be offering a lighter beer, than the state of Utah would probably best since there median ABV is the lowest in the US.

Within the writeup below you can find our analysis, regarding the beer & brewery data provided



```{r, echo=TRUE}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(dplyr)
library(na.tools)
library(sqldf)

#Set working directory
setwd("C:/Users/Pablo/Desktop/DS 6306 - Doing Data Science/Unit 7")

# Read CSV into R
beerData <- read.csv(file="Beers.csv", header=TRUE, sep=",")

breweryData <- read.csv(file="Breweries.csv", header=TRUE, sep=",")
```


#### 1. How many breweries are present in each state?
The image below displays the number of breweries present in each state.  As you will see, Colorado has highest the amount of breweries with 47.  Also, the following states are tied for having the the lowest amount of breweries with 1: West Virginia, South Dakota, North Dakota, & the District of Columbia. 


```{r, echo=TRUE, fig.height = 7, fig.width = 5}
#count of breweries per state
StateBrewCnt <- sqldf("select State, count(*) as Brewery_Count
                      from breweryData
                      group by State
                      order by Brewery_Count desc")

ggplot(data=StateBrewCnt, aes(x=reorder(State,Brewery_Count),y=Brewery_Count,fill=State)) + 
  geom_bar(stat="identity", show.legend = F) + 
  ggtitle("Count of Breweries Per State") +
  coord_flip() + 
  geom_text(data = StateBrewCnt, 
            aes(y = Brewery_Count, label = substr(Brewery_Count,0,5)), size = 3,
            vjust = .35, hjust=-.4) +
  xlab("State") +
  ylab("Brewery Count") +
  theme(
      panel.background = element_rect(fill = 'ivory1'),
      plot.title = element_text(hjust = 0.5),
      axis.text.y = element_text(size=8),
      axis.title.x = element_text(color="forestgreen", vjust=-0.35),
      axis.title.y = element_text(color="cadetblue" , vjust=0.35)   
        )

```


#### 2.   Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
To further our analysis, we have combined the brewery data set with the beer data set.  Within the image below, we have listed the first 6 rows & last 6 rows of the combined data.
```{r, echo=TRUE}

#Merging the two datasets based on Brewery ID
beersAndPubs <- merge(beerData,breweryData, by.x = "Brewery_id", by.y = "Brew_ID", all.x = TRUE)

names(beersAndPubs) <- c("Brewery_ID","Beer_Name","Beer_ID","ABV","IBU","Style","Ounces","Brewery_Name","City","State")

#Display first 6 rows and last 6 rows
head(beersAndPubs, 6)
tail(beersAndPubs,6)


```


#### 3.   Report the number of NA's in each column.
After continuing to review the combined data set, our group has identified a number of NAs (ie. missing values) within a few columns.  These will be reviewed and addressed within the next section.
```{r, echo=TRUE}

#Gather count of NA
na_count <-sapply(beersAndPubs, 
                  function(cnt) sum(length(which(is.na(cnt)))))

#Convert count vector to dataframe
na_df <- data.frame(na_count)

#Add new variable to list category names
na_df$categories <- c("Brewery_ID","Beer_Name","Beer_ID","ABV","IBU","Style","Ounces","Brewery_Name","City","State")

#plot the count data
ggplot(data=na_df, aes(x=categories, y=na_count)) + 
  geom_bar(stat="identity", fill="blue") +
  ggtitle("Count of NAs per Variable Category") +
  coord_flip() + 
  geom_text(data = na_df, 
            aes(y = na_count, label = na_count), size = 3,
            vjust = .35, hjust=-.4) +
  ylim(0,1050) +
  theme(
      panel.background = element_rect(fill = 'ivory1'),
      axis.text.y = element_text(size=8),
      axis.title.x = element_text(color="forestgreen", vjust=-0.35),
      axis.title.y = element_text(color="cadetblue" , vjust=0.35),
      plot.title = element_text(hjust = 0.5)
          ) +
  ylab("Count") +
  xlab("Categories")

```



#### 4.   Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
Before we began calculating the median ABV and IBU for each state, we applied an imputation for the NAs identified within the previous section.  Our method in filling out each NA, was to use the average ABV or IBU value calculated for each beer style.

The images below shows the median alcohol content (ABV) and median international bitterness unit (IBU) for each state.
```{r, echo=TRUE, fig.height = 7, fig.width = 5}

# This replaces the NA ABVs and IBUs in each state by the median of each State (not the national median)
beersAndPubs <- beersAndPubs %>% group_by(State) %>% mutate(ABV = na.median(ABV))
beersAndPubs <- beersAndPubs %>% group_by(State) %>% mutate(IBU = na.median(IBU))
# "All values of 'x' are missing (NA). Replacement value is 'NA'. Returning values unchanged" is thrown
# Whenever there is a state with no non-NA values to provide a median estimate with. To correct for this,
# these states will receive the nationwide median in the next step
beersAndPubs <- beersAndPubs %>% mutate(IBU = na.median(IBU)) %>% data.frame()
# Ensuring NAs were resolved (forced override if not - this was vetted with Proof of Concept):
beersAndPubs <- beersAndPubs %>% mutate(IBU = na.median(IBU)) %>% data.frame()

# creating ordered dataframe for overall medians for each state 
medianABVbyState <- beersAndPubs %>% group_by(State) %>% summarise(ABV = median(ABV)) %>% data.frame()
medianABVbyState <- data.frame(medianABVbyState[order(-medianABVbyState$ABV),])
medianIBUbyState <- beersAndPubs %>% group_by(State) %>% summarise(IBU = median(IBU)) %>% data.frame()
medianIBUbyState <- data.frame(medianIBUbyState[order(-medianIBUbyState$IBU),])

# ordered dataframes for collective ABVs and IBUs, by state
medianABVbyState
medianIBUbyState

###Steps below are to identify median ABV per style and update our model dataframe values accordingly###

#Separate IBU and Style into separate dataframe
abv_style_df <- beersAndPubs[,c("ABV","Style")]

#Calculate mean IBU per Style.  Also, rename column headers.
mn_abv <- aggregate(abv_style_df$ABV, by=list(abv_style_df$Style), FUN=median)
names(mn_abv) <- c("Style", "ABV")

#Plot Median ABV Per State
ggplot(data=medianABVbyState, aes(x=reorder(State,ABV),y=ABV,fill=State)) + 
  geom_bar(stat="identity", show.legend = F) + 
  ggtitle("Median ABV Per State") +
  coord_flip() + 
  geom_text(data = medianABVbyState, 
            aes(y = ABV, label = substr(ABV,0,5)), size = 3,
            vjust = .35, hjust=-.4) +
  ylim(0,0.09) +
  xlab("State") +
  theme(
      panel.background = element_rect(fill = 'ivory1'),
      plot.title = element_text(hjust = 0.5),
      axis.text.y = element_text(size=8),
      axis.title.x = element_text(color="forestgreen", vjust=-0.35),
      axis.title.y = element_text(color="cadetblue" , vjust=0.35)   
        )

#Plot Median IBU Per State
ggplot(data=medianIBUbyState, aes(x=reorder(State,IBU),y=IBU,fill=State)) + 
  geom_bar(stat="identity", show.legend = F) + 
  ggtitle("Median IBU Per State") +
  coord_flip() + 
  geom_text(data = medianIBUbyState, 
            aes(y = IBU, label = substr(IBU,0,5)), size = 3,
            vjust = .35, hjust=-.4) +
  xlab("State")+
  theme(
      panel.background = element_rect(fill = 'ivory1'),
      plot.title = element_text(hjust = 0.5),
      axis.text.y = element_text(size=8),
      axis.title.x = element_text(color="forestgreen", vjust=-0.35),
      axis.title.y = element_text(color="cadetblue" , vjust=0.35)   
        )
```




#### 5.   Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
Response to question an briefly exlain the purpose of the code chunk.
```{r, echo=TRUE}
# State with max median ABV (dataframe pre-ordered, so taking first row)
MaxABVstate <- (medianABVbyState[1,]) %>% data.frame()
print(MaxABVstate, row.names = F)

# State with max median IBU (dataframe pre-ordered, so taking first row)
MaxIBUstate <- (medianIBUbyState[1,]) %>% data.frame()
print(MaxIBUstate, row.names = F)

# State with the absolute max (singlemost - not median - strongest) ABV
print(beersAndPubs[which.max(beersAndPubs$ABV),c("State","ABV")], row.names = FALSE)

# State with the absolute max (singlemost - not median - bitter) IBU
print(beersAndPubs[which.max(beersAndPubs$IBU),c("State","IBU")], row.names = FALSE)

```




#### 6.   Summary statistics for the ABV variable.
Response to question an briefly exlain the purpose of the code chunk.
```{r, echo=TRUE}

# Probability density distribution for IBU
ggplot(na.omit(beersAndPubs), aes(x=IBU)) + geom_density(color="darkblue", fill="skyblue2") + theme(panel.background = element_rect(fill = 'ivory1'),plot.title = element_text(hjust = 0.5)) + ggtitle("Total Bitterness Distribution Across all States")

# Probability density distribution for ABV
ggplot(na.omit(beersAndPubs), aes(x=ABV)) + geom_density(color="darkblue", fill="skyblue2") + theme(panel.background = element_rect(fill = 'ivory1'),plot.title = element_text(hjust = 0.5)) + ggtitle("Total Alcohol by Volume Distribution Across all States")

### Boxplots
# ABV
ggplot(data=na.omit(beersAndPubs), aes(y=ABV,x=State)) + geom_boxplot(stat="boxplot", fill="skyblue2", color="darkblue") + theme(panel.background = element_rect(fill = 'ivory1'),plot.title = element_text(hjust = 0.5)) + ggtitle("Total ABV Across all U.S. States") + xlab("All States")
# IBU
ggplot(data=na.omit(beersAndPubs), aes(y=IBU,x=State)) + geom_boxplot(stat="boxplot", fill="skyblue2", color="darkblue") + theme(panel.background = element_rect(fill = 'ivory1'),plot.title = element_text(hjust = 0.5)) + ggtitle("Total Bitterness Across all States") + xlab("All States")

```



#### 7.   Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
Response to question an briefly exlain the purpose of the code chunk.
```{r, echo=TRUE}

#plot of IBU vs ABV with simple linear regression line
ggplot(beersAndPubs, aes(x=IBU, y=ABV, fill = IBU)) +
    geom_point(size=3, shape=21) +    
    geom_smooth(method=lm, se=FALSE) +
  theme(
        panel.background = element_rect(fill = 'ivory1'),
        plot.title = element_text(hjust = 0.5),
        legend.position = "none"
        ) +
  ggtitle("Total Alcohol by Volume vs. Bitterness Correlation")

```

